Bayesian exponential random graph modeling of KWL Burial network

### 2020-12-10 analysis/paper/paper.Rmd
edit discussion and work on conclusion

<!-- theoretical framework for the results-->
A striking finding in our results is the change in network properties of Kiwulan burials from a  more cohesive network with multiple subgroups (represented by higher transitivity and negative popularity) toward a more centralized network with  (lower transitivity and positive popularity) after the European presence during the 17th century. This supports our hypothesized models of chronological change from a more corporate society to a more networked society, as indicated in the grave goods and burial attributes. These observations are in line with theoretical assumptions of horizontal hierarchy, the corporate-network continuum that focuses on distinct strategies for achieving power as an alternative explanation to understand the emergence and level of social inequality [@Drennan2010; @Feinman2000; @Feinman2000political]. Compared to traditional models of vertical hierarchies that stress stratified social characteristics, the corporate-network model may be more relevant for understanding how a small scale society was organized, such as we see here in the pericolonial context of Kiwulan. A corporate-based society, such as the pre-European situation at Kiwulan, stresses shared power between individuals, communal  rituals, and social inequality, if any, would be associated with a group association [@Siegel1999]. In contrast, a more network-based society, indicated by the post-European data from Kiwulan, presents wealth accumulation through individual networks, resource control, and trade monopolization.

@@

<!-- other evidence and a bigger anthropological picture-->
We used burial data as a proxy to explore the social relations based on the argument that burials could be viewed as a social representation to understand social structures [@Binford1971; @Saxe1970]. We note that there could be an issue of disconnection between the role of the deceased in a burial context and their real life, which should be reduced by examining evidence from the residential context [@Chapman2003]. Previous studies of trade ornaments from the residential area of Kiwulan suggests an uneven spatial distribution when the Europeans were active in northern Taiwan that hints at an increased social differentiation at Kiwulan in a pericolonial context [@LWandBM2020ornament]. This is consistent with the result of a more centralized burial network after the presence of the Europeans, which supports the assumption of social role representation from burials in this case. Since the burial goods used for making ties were treated as prestige goods in local Indigenous culture [@Borao2009; @LiandWu2006], the interrelations represented by the flow of goods could reflect the process of social changes at Kiwulan in a pericolonial context. After the arrival of Europeans, some individuals may be differentiated or privileged to possess more distinctive goods through becoming the trade partners with the Europeans. According to historical records, the Dutch would give local Indigenous leaders symbolic items or trade goods to assert their authority and consolidate the relationship with local people [@Kang2016, pp137-162]. Uneven distribution of goods in burials we observed could result from differential access to exotic trade relative to the European colonization and gradually led to increased social differentiation. The involvement in long-distance exchange often leads to changes in political-economic strategies to a network mode, where ambitious individuals are able to build their factions and personal influence through the distribution of high value goods [@Feinman2000; ].

<!-- methodological implication-->
The ERGM network analysis within a Bayesian framework provides a robust way to examine relations between burials based on grave goods and attributes associated with demography and mortuary practices. The adoption of Bayesian inference on ERGM can solve the issues of computational tractability and degeneracy in parameter estimation in ERGM [@Caimo2017]. In addition, a Bayesian framework reduced the effects of small sample size or missing data commonly present in archaeological data by incorporating prior information and MCMC estimations. We demonstrated a plausible and robust method to test the hypothesized model of chronological changes in burial structures with probabilistic interpretation. Although some network statistics of covariants do not show the pattern as we assumed, the Bayesian approach is an efficient and effective method for evaluating the endogenous network effects relative to the overall structure of burials. This structure can be viewed as a representation of a given society.

<!-- more discussion of nodal attributes -->
Age and sex attributes serve as indicators for distinguishing ascribed or achieved status in a society [citation].

# Conclusion
It should be noted that corporate/network spectrum represents a different degree of hierarchical complexity instead of a clear division. 
This might lead to accumulation of rare prestige goods and competition between individuals that Kiwulan gradually moves to a more network-based society.

>>>>

<!-- theoretical framework for the results-->
A striking finding in our results is the change in network properties of Kiwulan burials from a more cohesive network with multiple subgroups (represented by higher transitivity and negative popularity) toward a more centralized network with concentration in a limited burials (lower transitivity and positive popularity) after the European presence during the 17th century. This supports our hypothesized models of chronological change from a more corporate society to a more networked society, as indicated in the grave goods and burial attributes. These observations are in line with theoretical assumptions of horizontal hierarchy, the corporate-network continuum that focuses on distinct strategies for achieving power as an alternative explanation to understand the emergence and level of social inequality [@Drennan2010; @Feinman2000; @Feinman2000political]. Compared to traditional models of vertical hierarchies that stress stratified social characteristics, the corporate-network model may be more relevant for understanding how a small scale society was organized, such as we see here in the pericolonial context of Kiwulan. A corporate-based society, such as the pre-European situation at Kiwulan, stresses shared power between individuals, communal  rituals, and social inequality, if any, would be associated with a group association [@Siegel1999]. In contrast, a more network-based society, indicated by the post-European data from Kiwulan, presents wealth accumulation through individual networks, prestige good system manipulation, and trade monopolization [@Blanton1996]. It should be noted that corporate/network continuum represents a dynamic process with different degree of hierarchical complexity instead of a static ideal-type stage.

@@

<!-- other evidence -->
We used burial data as a proxy to explore the social relations based on the argument that burials could be viewed as a social representation to understand social structures [@Binford1971; @Saxe1970]. We note that there could be an issue of disconnection between the role of the deceased in a burial context and their real life, which should be reduced by examining evidence from the residential context [@Chapman2003]. Previous studies of trade ornaments from the residential area of Kiwulan suggests an uneven spatial distribution when the Europeans were active in northern Taiwan that hints at an increased social differentiation at Kiwulan in a pericolonial context [@LWandBM2020ornament]. This is consistent with the result of a more centralized burial network after the presence of the Europeans, which supports the assumption of social role representation from burials in this case. Since the burial goods used for making ties were treated as prestige goods or status items in local Indigenous culture [@Borao2009; @LiandWu2006], the interrelations represented by the flow of goods could reflect the process of social changes at Kiwulan in a pericolonial context.

<!--fit in a big anthropological picture in a periclonial context -->
The involvement in long-distance exchange is often associated with changes in political-economic strategies to a network mode, where ambitious individuals are able to build their local power base and personal influence through the distribution of high-value goods [@Blanton1996; @Feinman2000; @Klehm2017]. Northeastern Taiwan was involved in a regional exchange in East Asia since the 14th century and long-distance trade during the European presence in the 17th century [@Chen2005; @Wang2007]. Despite relatively weak control, northeastern Taiwan was incorporated into colonial territories in 1647 after the Dutch conquests []. The Dutch would give local Indigenous leaders symbolic items or trade goods to assert their political authority and consolidate their relationship with Indigenous societies [@Kang2016]. At Kiwulan, after the arrival of Europeans, some individuals may be differentiated or privileged to possess more distinctive goods through becoming the trade partners with the Europeans. Changes in interburial networks we observed may result from differential access to exotic trade relative to the European colonization and gradually led to an increased social differentiation. The image of European power could be embedded into prestige objects pre-existing in Indigenous culture where their values were amplified with concepts of wealth or power, since good values are contextually mutable and entangled in historical processes [@Aswani2003; @Thomas2009]. This may lead to ambitious individuals to compete with each other and accumulate trade goods. It demonstrates the agency of local Indigenous societies to incorporate and manipulate trade items in a precolonial context, gradually resulting in social changes from a more corporate to a more network strategy.

# Conclusion

<!-- summary of the research, contribution and methodological implication-->
In this paper, we presented a naval approach for studying burial relations using ERGM network analysis within a Bayesian framework to rigorously examined social changes in a European percicolonial context in northeastern Taiwan. We tested the hypotheses of chronological changes in interburial networks at Kiwulan with the evaluation of both endogenous and exogenous network effects. Our results provide a meaningful explanation for the relationship between burials that share similar prestige goods. Before European contact in the 17th century, the network properties has a tendency of more closed burial relationships with ritual elements as the key formation mechanism. After the European arrival, the network has a tendency of centralization relative to the rarity of goods. This aliens with changes in social complexity from a more corporate to a more network mode society that stress different strategies for achieving power and different degrees of social inequality. Based on the burial data, along with historical records and ethnography, we are able to understand the processes of changes in Indigenous societies in the periphery of European colonial bases. This study demonstrates not only the pericolonial effect on local Indigenous societies that have not well understood yet, but also the agency of local people to manipulate colonial power instead of passive recipients of trade goods.

This case study highlights the methodological improvement that Bayesian inference on ERGM can inform and enhance studies in rational data in archaeology. A Bayesian framework reduced the effects of small sample size or missing data commonly present in archaeological data by incorporating prior information and MCMC estimations. Moreover, the interrelations indicated by the probabilistic inference provide robust results, allowing in-depth discussion for the dynamic processes of relationship formation and their underlying factors to reflect the historical trajectories of sociocultural phenomena. This approach can be applied to a wide range of relational data in archaeology to examine the formation of relationships and give an insight into the study of social structure in general.
